---
title: "Budweiser Case Study"
output:
  html_document:
    toc: true
    toc_depth: 2
author: "Kevin Boyd & Scott Frazier"
date: "10/6/2021"
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
Our goal was to perform an EDA by identifying errors and recognizing patterns that are present in the beers and breweries data sets given to us by Budweiser. Much of the work consisted of creating visual aids to show insights from the data. We addressed the relationship between alcohol content and the bitterness using multiple plots and methods. We also utilized a k-NN model to address the relationship between ALE and IPA. As a result this EDA shows how some of the most important components of beer relate to each other.


# Load Libraries
```{r, message=F}
library(tidyverse)
library(RColorBrewer)
library(cowplot)
library(class)
library(caret)
library(e1071)
library(stringr)
library(maps)
library(mapproj)
```


# Load in the data
```{r}
beers <- read.csv("beers.csv", header = T)
breweries <- read.csv("breweries.csv", header = T)
```


# Look at the data
```{r}
str(beers)
str(breweries)
head(beers)
head(breweries)
```


# Number 1 - Number of Breweries in Each State
```{r}
#ordered plot showing counts of breweries per state
brewbystate <- breweries %>%
  group_by(State) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=reorder(State,(-count)), y = count)) +
  geom_bar(stat = "identity",fill="blue") +
  theme(axis.text.x=element_text(angle=90)) +
  ggtitle("Number of Breweries by State") +
  xlab("State") +
  ylab("Count")
brewbystate + theme(axis.text.x=element_text(vjust=0.5))
```


# Number 2 - Merge the Datasets
```{r}
#merge data sets on Brewery_id
names(breweries)[1] <- "Brewery_id"
df1 <- merge(beers, breweries, by = "Brewery_id", all = T)
names(df1)[2] <- "Beer"
names(df1)[8] <- "Brewery"
str(df1)
head(df1)
tail(df1)
```


# Number 3 - Missing Values
```{r}
#number of NAs in each data set
Beers_NAs <- sum(is.na(beers))
Brewery_NAs <- sum(is.na(breweries))
NA.df <- data.frame(Measure = c("Beers Data", "Breweries Data"), 
                    Value = c(Beers_NAs,Brewery_NAs))
NA.dataset <- ggplot(NA.df, aes(x=Measure, y=Value)) +
  geom_bar(stat = "identity", fill = "red3") +
  ggtitle("Missing Data in each Dataset") +
  xlab("Dataset") +
  ylab("Number of NAs")

#which rows are the NAs coming from?
ABV_NAs <- sum(is.na(beers$ABV))
IBU_NAs <- sum(is.na(beers$IBU))
DataLength <- NROW(beers$ABV)

#make data frame from number of NAs
NA.df1 <- data.frame(Measure = c("ABV","IBU"),
                     Value = c(ABV_NAs,IBU_NAs))

#plot the NA values from each column
NA.column <- ggplot(NA.df1, aes(x=Measure, y=Value)) +
  geom_bar(stat = "identity", fill = "purple2") +
  ggtitle("Missing Data for ABV and IBU") +
  ylab("Number of NAs") +
  xlab("Column in Dataset")

#make a pie chart for ABV
NA.df2 <- data.frame(Measure = c("ABV","ABV"), 
                     Data = c("Missing","Total"),
                     Value = c(ABV_NAs,(DataLength-ABV_NAs)))

bp1 <- ggplot(NA.df2, aes(x="", y=Value, fill=Data)) +
  geom_bar(width = 1, stat = "identity") + xlab("") + ylab("") + ggtitle("Missing ABV Data")
pie1 <- bp1 + coord_polar("y", start=0)
pie1

#make pie chart for IBU
NA.df3 <- data.frame(Measure = c("IBU","IBU"), 
                     Data = c("Missing","Total"),
                     Value = c(IBU_NAs,(DataLength - IBU_NAs)))

bp2 <- ggplot(NA.df3, aes(x="", y=Value, fill=Data)) +
  geom_bar(width = 1, stat = "identity") + xlab("") + ylab("") + ggtitle("Missing IBU Data")
pie2 <- bp2 + coord_polar("y", start=0)
pie2

#put all the plots together
NAplots <- plot_grid(NA.dataset,NA.column,pie1,pie2, labels = c("A","B","C","D"))
NAplots
```


# Number 4a - Median ABV by State
```{r}
#plot showing the median ABV value per state
ABV_Values <- df1 %>% group_by(State) %>%
  summarize(medianABV = median(ABV, na.rm = T)) %>%
  ggplot(aes(x=reorder(State,(-medianABV)), y=medianABV)) +
  geom_bar(stat = "identity", fill="#377eb8") +
  theme(axis.text.x=element_text(angle=90)) +
  ggtitle("Median ABV Values by State") +
  xlab("State") +
  ylab("Median ABV Value")
ABV_Values + theme(axis.text.x=element_text(vjust=0.5))
```


# Number 4b - Median IBU by State
```{r}
#plot showing the median IBU value per state
IBU_Values <- df1 %>% group_by(State) %>%
  summarize(medianIBU = median(IBU, na.rm = T)) %>%
  ggplot(aes(x=reorder(State,(-medianIBU)), y=medianIBU)) +
  geom_bar(stat = "identity", fill="#4daf4a") +
  theme(axis.text.x=element_text(angle=90)) +
  ggtitle("Median IBU Values by State") +
  xlab("State") +
  ylab("Median IBU Value")
IBU_Values + theme(axis.text.x=element_text(vjust=0.5))
```


# Number 5a - Maximum ABV and IBU
```{r}
#make data frames with top 5 ABV and IBU values
df2 <- df1[order(df1$ABV,decreasing = T),]
df2 <- df2[c(1,2,3,5,6),c(4,10)]
df3 <- df1[order(df1$IBU,decreasing = T),]
df3 <- df3[c(1:5),c(5,10)]
```

# Number 5b - Top 5 States with Highest IBU
**Oregon has the highest IBU beer**
```{r}
#plot the top 5 states for IBU
ggplot(df3, aes(x = reorder(State, -IBU), y = IBU)) + 
  geom_bar(stat = "identity", fill = "#4daf4a") +
  ylim(0,150) +
  ggtitle("Top 5 States IBU") +
  xlab("State") +
  geom_text(aes(label=IBU), vjust=-1) +
  ylab("IBU (International Bitterness Units)")
```

# Number 5c - Top 5 States with Highest ABV
**Colorado has the highest ABV beer**
```{r}
#plot the top states for ABV
ggplot(df2, aes(x = reorder(State, -ABV), y = ABV)) + 
  geom_bar(stat = "identity", fill= "#377eb8") +
  ylim(0,0.15) +
  ggtitle("Top 5 States ABV") +
  xlab("State") +
  geom_text(aes(label=ABV), vjust=-1) +
  ylab("ABV (Alcohol by Volume)")
```


# Number 6 - Distribution of ABV
**The evidence suggests the distribution of ABV is right skewed with a median value of 0.056 with a mean of 0.0598. There are a number of values greater than 0.1 that make it appear to be right skewed.**
```{r}
#Summary stats of ABV
summary(df1$ABV)

#Distributions
#Histogram of ABV in df1
df1 %>% ggplot(aes(x=ABV)) + geom_histogram(fill = "#377eb8") + ggtitle("Different ABV Values of Beer") + labs(x="ABV (Alcohol by Volume)", y="Frequency")

#boxplot
df1 %>% ggplot(aes(x=ABV)) + geom_boxplot(fill = "#377eb8") + ggtitle("Different ABV Values of Beer") + labs(x="ABV (Alcohol by Volume)", y="") + coord_flip()


#geom_density
ABV_vizz = ggplot(df1, aes(x = ABV))

ABV_vizz + geom_density(fill = "#377eb8", alpha=0.6) +
  geom_vline(aes(xintercept = mean(ABV)), 
             linetype = "dashed", size = 0.6,
             color = "#FC4E07") +
  ylab("Frequency") +
  ggtitle("Different ABV Values of Beer") +
  xlab("ABV (Alcohol by Volume)")
```


# Number 7 - Relationship between ABV and IBU
**There is evidence of a positive relationship between ABV and IBU. We also notice that a majority of the data points are at 0.05 ABV and 35 IBU**
```{r}
ggplot(df1, aes(x=ABV, y=IBU)) +
  geom_point() +
  geom_jitter() +
  geom_smooth() +
  ggtitle("Relationship Between ABV vs IBU")
```


# Number 8a - k-NN Model
```{r}
#Make new column with ALE, IPA, or Other as identifier
df1$Type = ifelse(grepl("\\bale\\b", df1$Style, ignore.case = T), "ALE",
                  ifelse(grepl("\\bipa\\b", df1$Style, ignore.case = T), "IPA", "Other"))
head(df1)

#Make new data frame with only Ale beers
aleIPA <- df1 %>% filter(df1$Type == "ALE" | df1$Type == "IPA")
head(aleIPA)

#Make training and test data frames
splitPerc <- .7
trainIndices <- sample(1:dim(aleIPA)[1],round(splitPerc * dim(aleIPA)[1]))
dfTrain <- aleIPA[trainIndices,]
dfTest <- aleIPA[-trainIndices,]

#Omit the NAs
dfTrain <- na.omit(dfTrain)
dfTest <- na.omit(dfTest)

#knn model
classifications <- knn(dfTrain[,4:5], dfTest[,4:5], dfTrain$Type, k = 10, prob = TRUE)
table(dfTest$Type,classifications)
confusionMatrix(table(dfTest$Type,classifications))
```

# Number 8b - Find Best k Value
```{r}
#Loop for many k and the average of many training / test partition
iterations = 100
numks = 50

masterAcc = matrix(nrow = iterations, ncol = numks)

for(j in 1:iterations)
{
  accs = data.frame(accuracy = numeric(100), k = numeric(100))
  trainIndices = sample(1:dim(aleIPA)[1],round(splitPerc * dim(aleIPA)[1]))
  train = aleIPA[trainIndices,]
  test = aleIPA[-trainIndices,]
  train = na.omit(train)
  test = na.omit(test)
  for(i in 1:numks)
  {
    classifications = knn(train[,c(4,5)],test[,c(4,5)],train$Type, prob = TRUE, k = i)
    table(classifications,test$Type)
    CM = confusionMatrix(table(classifications,test$Type))
    masterAcc[j,i] = CM$overall[1]
  }
  
}

MeanAcc = colMeans(masterAcc)

plot(seq(1,numks,1),MeanAcc, type = "l", xlab = "k", ylab = "Mean Accuracy", main = "Mean Accuracy of k Values")
```

# Number 8c - k-NN Model with Optimized k
```{r}
#k-NN model using best value of k 
classifications4 <- knn(dfTrain[,4:5], dfTest[,4:5], dfTrain$Type, k = 6, prob = TRUE)
table(dfTest$Type,classifications4)
confusionMatrix(table(dfTest$Type,classifications4))
```


# Number 8d - Test ABV & IBU Contribution to Model Accuracy
**Both ABV and IBU allowed high accuracy in our model predicting if a beer is considered an IPA, but IBU was the best measure in the model.**
```{r}
#knn internal model for ABV and IBU
classifications1 <- knn.cv(dfTrain[,4:5],dfTrain$Type, k = 10)
confusionMatrix(table(classifications1,dfTrain$Type))

#knn model for ABV
classifications2 <- knn.cv(dfTrain[,4],dfTrain$Type, k = 10)
confusionMatrix(table(classifications2,dfTrain$Type))

#knn model for IBU
classifications3 <- knn.cv(dfTrain[,5],dfTrain$Type, k = 10)
confusionMatrix(table(classifications3,dfTrain$Type))
```


# Number 9a - Interesting Relationships
```{r}
# make new column with beer identifiers and make into a factor
df1$New <- ifelse(grepl("\\bale\\b", df1$Style, ignore.case = T), "ALE",
                  ifelse(grepl("\\bipa\\b", df1$Style, ignore.case = T), "IPA",
                         ifelse(grepl("\\bWheat\\b", df1$Style, ignore.case = T), "Wheat",
                                ifelse(grepl("\\bAmber\\b", df1$Style, ignore.case = T), "Amber",
                                       ifelse(grepl("\\bStout\\b", df1$ Style, ignore.case = T), "Stout",
                                              ifelse(grepl("\\bLager\\b", df1$Style, ignore.case = T), "Lager",
                                                     ifelse(grepl("\\bPorter\\b", df1$Style, ignore.case = T), "Porter",
                                                            ifelse(grepl("\\bPilsner\\b", df1$Style, ignore.case = T), "Pilsner",
                                                                   "Other"))))))))
head(df1)
df1$New <- as.factor(df1$New)
str(df1)
```

```{r}
#plot each beer ABV and IBU individually
ggplot(df1, aes(x=ABV, y=IBU, color=New)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(color="black") +
  ggtitle("Relationship Between ABV vs IBU") +
  labs(color="Type of Beer") +
  facet_wrap(~New)
```

```{r}
#boxplot for ABV of each beer
p1 <- ggplot(df1, aes(x=New, y=ABV, fill=New)) +
  geom_boxplot() +
  ggtitle("Relationship Between Kind of Beer and ABV") +
  xlab("Type of Beer") + 
  theme(legend.position="none")
p1
```

```{r}
#boxplot for IBU of each beer
p2 <- ggplot(df1, aes(x=New, y=IBU, fill=New)) +
  geom_boxplot() +
  ggtitle("Relationship Between Kind of Beer and IBU") +
  xlab("Type of Beer") + 
  theme(legend.position="none")
p2
```


# Number 9b - Median ABV and IBU by State
```{r}
#make the data frame to plot
df3.ABV <- df1 %>% group_by(State) %>%
  summarize(medianABV = median(ABV, na.rm = T))

df3.IBU <- df1 %>% group_by(State) %>%
  summarize(medianIBU = median(IBU, na.rm = T))

df4 <- cbind(df3.ABV,df3.IBU)
df4 <- df4[,-3]
colnames(df4)[1] <- "abb"
df4 <- df4[-8,]
lookup = data.frame(abb = state.abb, State = state.name)
df5 <- lookup[order(lookup$abb),]
df6 <- cbind(df4,df5)
df6$region <- tolower(df6$State)
df7 <- df6[,-c(1,4,5)]

#make map df
states <- map_data("state")
map.df1 <- merge(states,df7, by="region", all.x=T)
map.df1 <- map.df1[order(map.df1$order),]
```

```{r}
#ABV heat plot
ggplot(map.df1, aes(x=long,y=lat,group=group)) +
  geom_polygon(aes(fill=medianABV)) +
  geom_path() + 
  scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey90") +
  ggtitle("Median ABV by State") +
  labs(fill="Median ABV") +
  coord_map()
```

```{r}
#IBU heat plot
ggplot(map.df1, aes(x=long,y=lat,group=group)) +
  geom_polygon(aes(fill=medianIBU)) +
  geom_path() + 
  scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey90") +
  ggtitle("Median IBU by State") +
  labs(fill="Median IBU") +
  coord_map()
```

# Conclusion
**Based on our visualizations and insights, we concluded that there does appear to be evidence of a positive relationship between ABV and IBU. We were able to accuratly identify if a beer was an ale or an IPA based specifically on those metrics. We also observed similar distributions of ABV and IBU between different subgroups of beer.**