---
title: "Budweiser Case Study"
author: "Kevin Boyd & Scott Frazier"
date: "10/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Budweiser Case Study


### Load Libraries
```{r, message=F}
library(tidyverse)
library(RColorBrewer)
library(cowplot)
```


### Load in the data
```{r}
beers <- read.csv("beers.csv", header = T)
breweries <- read.csv("breweries.csv", header = T)
```


### Look at the data
```{r}
str(beers)
str(breweries)
head(beers)
head(breweries)
```


### How many breweries in each state
Number 1
```{r}
#ordered plot
breweries %>%
  group_by(State) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=reorder(State,(-count)), y = count)) +
  geom_bar(stat = "identity",fill="blue") +
  theme(axis.text.x=element_text(angle=90)) +
  ggtitle("Number of Breweries by State") +
  xlab("State") +
  ylab("Count")
```


### Merge datasets
Number 2
```{r}
#merge data sets on Brewery_id
names(breweries)[1] <- "Brewery_id"
df1 <- merge(beers, breweries, by = "Brewery_id", all = T)
names(df1)[2] <- "Beer"
names(df1)[8] <- "Brewery"
str(df1)
head(df1)
tail(df1)
```


### How many NAs
Number 3
```{r}
#number of NAs in each data set
Beers_NAs <- sum(is.na(beers))
Brewery_NAs <- sum(is.na(breweries))
NA.df <- data.frame(Measure = c("Beers Data", "Breweries Data"), 
                                Value = c(Beers_NAs,Brewery_NAs))
NA.dataset <- ggplot(NA.df, aes(x=Measure, y=Value)) +
  geom_bar(stat = "identity", fill = "red3") +
  ggtitle("Missing Data in each Dataset") +
  xlab("Dataset") +
  ylab("Number of NAs")

#which rows are the NAs coming from?
ABV_NAs <- sum(is.na(beers$ABV))
IBU_NAs <- sum(is.na(beers$IBU))
DataLength <- NROW(beers$ABV)

#make data frame from number of NAs
NA.df1 <- data.frame(Measure = c("ABV","IBU"),
                     Value = c(ABV_NAs,IBU_NAs))

#plot the NA values from each column
NA.column <- ggplot(NA.df1, aes(x=Measure, y=Value)) +
  geom_bar(stat = "identity", fill = "purple2") +
  ggtitle("Missing Data for ABV and IBU") +
  ylab("Number of NAs") +
  xlab("Column in Dataset")

#make a pie chart for ABV
NA.df2 <- data.frame(Measure = c("ABV","ABV"), 
                     Data = c("Missing","Total"),
                     Value = c(ABV_NAs,(DataLength-ABV_NAs)))

bp1 <- ggplot(NA.df2, aes(x="", y=Value, fill=Data)) +
geom_bar(width = 1, stat = "identity") + xlab("") + ylab("") + ggtitle("Missing ABV Data")
pie1 <- bp1 + coord_polar("y", start=0)
pie1

#make pie chart for IBU
NA.df3 <- data.frame(Measure = c("IBU","IBU"), 
                     Data = c("Missing","Total"),
                     Value = c(IBU_NAs,(DataLength - IBU_NAs)))

bp2 <- ggplot(NA.df3, aes(x="", y=Value, fill=Data)) +
geom_bar(width = 1, stat = "identity") + xlab("") + ylab("") + ggtitle("Missing IBU Data")
pie2 <- bp2 + coord_polar("y", start=0)
pie2

#put all the plots together
NAplots <- plot_grid(NA.dataset,NA.column,pie1,pie2, labels = c("A","B","C","D"))
NAplots
```


### Median ABV by State
Number 4
```{r}
ABV_Values <- df1 %>% group_by(State) %>%
  summarize(medianABV = median(ABV, na.rm = T)) %>%
  ggplot(aes(x=reorder(State,(-medianABV)), y=medianABV)) +
  geom_bar(stat = "identity", fill="#377eb8") +
  theme(axis.text.x=element_text(angle=90)) +
  ggtitle("Median ABV Values by State") +
  xlab("State") +
  ylab("Median ABV Value")
ABV_Values
```


### Median IBU by State
```{r}
IBU_Values <- df1 %>% group_by(State) %>%
  summarize(medianIBU = median(IBU, na.rm = T)) %>%
  ggplot(aes(x=reorder(State,(-medianIBU)), y=medianIBU)) +
  geom_bar(stat = "identity", fill="#4daf4a") +
  theme(axis.text.x=element_text(angle=90)) +
  ggtitle("Median IBU Values by State") +
  xlab("State") +
  ylab("Median IBU Value")
IBU_Values
```


### Maximum ABV and IBU
#### Number 5
##### Colorado has the highest ABV beer and Oregon has the highest IBU beer
```{r}
#make data frames with top 5 
df2 <- df1[order(df1$ABV,decreasing = T),]
df2 <- df2[c(1,2,3,5,6),c(4,10)]
df3 <- df1[order(df1$IBU,decreasing = T),]
df3 <- df3[c(1:5),c(5,10)]
```

### Top 5 States with Highest IBU
```{r}
#plot the top 5 states for IBU
ggplot(df3, aes(x = reorder(State, -IBU), y = IBU)) + 
  geom_bar(stat = "identity", fill = "#4daf4a") +
  ylim(0,150) +
  ggtitle("Top 5 States IBU") +
  xlab("State") +
  geom_text(aes(label=IBU), vjust=-1) +
  ylab("IBU (International Bitterness Units)")
```

### Top 5 States with Highest ABV
```{r}
#plot the top states for ABV
ggplot(df2, aes(x = reorder(State, -ABV), y = ABV)) + 
  geom_bar(stat = "identity", fill= "#377eb8") +
  ylim(0,0.15) +
  ggtitle("Top 5 States ABV") +
  xlab("State") +
  geom_text(aes(label=ABV), vjust=-1) +
  ylab("ABV (Alcohol by Volume)")
```


### Distribution of ABV
#### Number 6
#####  The distribution of ABV is right skewed with a median value of 0.056 with a mean of 0.0598. There are a number of values greater than 0.1 that make it right skewed.
```{r}
#Summary stats of ABV
summary(df1$ABV)

###Distributions
#Histogram of ABV in df1 - looks decent
df1 %>% ggplot(aes(x=ABV)) + geom_histogram(fill = "#377eb8") + ggtitle("Different ABV Values of Beer") + labs(x="ABV (Alcohol by Volume)", y="Frequency")

#boxplot - might need some work
df1 %>% ggplot(aes(x=ABV)) + geom_boxplot(fill = "#377eb8") + ggtitle("Different ABV Values of Beer") + labs(x="ABV (Alcohol by Volume)", y="") + coord_flip()


#geom_density - looks decent
ABV_vizz = ggplot(df1, aes(x = ABV))

ABV_vizz + geom_density(fill = "#377eb8", alpha=0.6) +
  geom_vline(aes(xintercept = mean(ABV)), 
             linetype = "dashed", size = 0.6,
             color = "#FC4E07") +
  ylab("Frequency") +
  ggtitle("Different ABV Values of Beer") +
  xlab("ABV (Alcohol by Volume)")
```


### Relationship between ABV and IBU
#### Number 7
##### There appears to be a positive relationship between ABV and IBU. As the ABV increases the IBU also tends to increase. We also notice that a majority of the data points are at 0.05 ABV and 35 IBU
```{r}
ggplot(df1, aes(x=ABV, y=IBU)) +
  geom_point() +
  geom_jitter() +
  geom_smooth() +
  ggtitle("Relationship Between ABV vs IBU")
```





